### 2017-11-28 submission/manuscript.Rmd
add imp otu and new figs

***A Majority of Studies are Underpowered for Detecting Small Effect Size Differences:***
Based on the previous observations we then assessed whether the studies that we included are realistically powered to identify small, medium, and large scale differences between case and control. When assessing the power of each study at different effect sizes the majority of studies for both adenoma and carcinoma have an 80% power to detect a 30% difference [Figure 6A & B]. No single study that was analyzed had the standard 80% power to detect an effect size difference that was equal to or below 10% [Figure 6A & B].  In order to achieve adequate power for small effect sizes, studies would need to recruit over 1000 individuals for each arm [Figure 6C].

@@

@@

**Figure 6: Power and Effect Size Analysis of Studies Included.** A) Power based on effect size for studies with adenoma individuals. B) Power based on effect size for studies with carcinoma individuals. C) The estimated sample number needed for each arm of each study to detect an effect size of 1-30%. The dotted red lines in A) and B) represent a power of 0.8.

@@

>>>>

In stool the most common genera in the genus based models belonged mostly to resident genera [Figure 6A & B]. This included genera such as *Ruminococcus*, *Bacteroides*, and *Roseburia*. With respect to the CRC-associated genera *Fusobacterium* was the only genus present in adenoma while all four were present in carcinoma [Figure 6A & B]. However, none of these CRC-associated genera were present in the majority of studies. For the adenoma OTU models, OTUs that classified as  *Ruminococcaceae* or *Roseburia* were present in the top 10 OTUs for the vast majority of studies [Figure 6C]. Although for the carcinoma OTU models *Ruminococcaceae* was in the top 10 for many studies, *Bacteroides* was present in the overwhelming majority of the carcinoma OTU stool models [Figure 6D].

Conversely to the stool models, both genera and OTU based models in tissue had the vast majority of their top 10 occur in a study specific manner [Figure S7]. *Fusobacterium* and *Fusobacteriaceae* show up more often in the top 10 for matched tissue samples but are not present in the top 10 or much lower for unmatched tissue [Figure S7B-C & S7E-F]. There appears to be very little overlap in the top 10 most important variables between stool and tissue for both adenoma and carcinoma [Figure 6 & S7].

***A Majority of Studies are Underpowered for Detecting Small Effect Size Differences:***
Based on the previous observations we then assessed whether the studies that we included are realistically powered to identify small, medium, and large scale differences between case and control. When assessing the power of each study at different effect sizes the majority of studies for both adenoma and carcinoma have an 80% power to detect a 30% difference [Figure 7A & B]. No single study that was analyzed had the standard 80% power to detect an effect size difference that was equal to or below 10% [Figure 7A & B].  In order to achieve adequate power for small effect sizes, studies would need to recruit over 1000 individuals for each arm [Figure 7C].

@@

***Assessing Important Random Forest Model Variables:***
The genus based models collected the top 10 most important variables from each training set and assessed how many times that genera showed up in the top 10 across each study. The OTU based models recorded the medians for each OTU across 100 different 80/20 splits of the data for each study. The lowest classification for each OTU was obtained using the RDP database and the number of times the specific classification occured in the top 10 across studies was recorded. For the adenoma tissue genus and OTU models there was only one matched and unmatched study and these results were grouped together for the counting of the top 10.

@@

**Figure 6: Most Common Genera Across Full Community Stool Study Models.** A) Common genera in the top 10 for adenoma Random Forest genus models. B) Common genera in the top 10 for carcinoma Random Forest genus models. C) Common genera in the top 10 for adenoma Random Forest OTU models. D) Common genera in the top 10 for carcinoma Random Forest OTU models.

**Figure 7: Power and Effect Size Analysis of Studies Included.** A) Power based on effect size for studies with adenoma individuals. B) Power based on effect size for studies with carcinoma individuals. C) The estimated sample number needed for each arm of each study to detect an effect size of 1-30%. The dotted red lines in A) and B) represent a power of 0.8.

@@

**Figure S7: Most Common Genera Across Full Community Tissue Study Models.** A) Common genera in the top 10 for adenoma Random Forest genus models. B) Common genera in the top 10 for unmatched carcinoma Random Forest genus models. B) Common genera in the top 10 for matched carcinoma Random Forest genus models. D) Common genera in the top 10 for adenoma Random Forest OTU models. E) Common genera in the top 10 for unmatched carcinoma Random Forest OTU models. F) Common genera in the top 10 for matched carcinoma Random Forest OTU models.

### 2017-11-28 submission/manuscript.tex
add imp otu and new figs

\textbf{\emph{A Majority of Studies are Underpowered for Detecting Small
Effect Size Differences:}} Based on the previous observations we then
assessed whether the studies that we included are realistically powered
to identify small, medium, and large scale differences between case and
control. When assessing the power of each study at different effect
sizes the majority of studies for both adenoma and carcinoma have an
80\% power to detect a 30\% difference {[}Figure 6A \& B{]}. No single
study that was analyzed had the standard 80\% power to detect an effect
size difference that was equal to or below 10\% {[}Figure 6A \& B{]}. In
order to achieve adequate power for small effect sizes, studies would
need to recruit over 1000 individuals for each arm {[}Figure 6C{]}.

@@

@@

\textbf{Figure 6: Power and Effect Size Analysis of Studies Included.}
A) Power based on effect size for studies with adenoma individuals. B)
Power based on effect size for studies with carcinoma individuals. C)
The estimated sample number needed for each arm of each study to detect
an effect size of 1-30\%. The dotted red lines in A) and B) represent a
power of 0.8.

@@

>>>>

In stool the most common genera in the genus based models belonged
mostly to resident genera {[}Figure 6A \& B{]}. This included genera
such as \emph{Ruminococcus}, \emph{Bacteroides}, and \emph{Roseburia}.
With respect to the CRC-associated genera \emph{Fusobacterium} was the
only genus present in adenoma while all four were present in carcinoma
{[}Figure 6A \& B{]}. However, none of these CRC-associated genera were
present in the majority of studies. For the adenoma OTU models, OTUs
that classified as \emph{Ruminococcaceae} or \emph{Roseburia} were
present in the top 10 OTUs for the vast majority of studies {[}Figure
6C{]}. Although for the carcinoma OTU models \emph{Ruminococcaceae} was
in the top 10 for many studies, \emph{Bacteroides} was present in the
overwhelming majority of the carcinoma OTU stool models {[}Figure 6D{]}.

Conversely to the stool models, both genera and OTU based models in
tissue had the vast majority of their top 10 occur in a study specific
manner {[}Figure S7{]}. \emph{Fusobacterium} and \emph{Fusobacteriaceae}
show up more often in the top 10 for matched tissue samples but are not
present in the top 10 or much lower for unmatched tissue {[}Figure S7B-C
\& S7E-F{]}. There appears to be very little overlap in the top 10 most
important variables between stool and tissue for both adenoma and
carcinoma {[}Figure 6 \& S7{]}.

\textbf{\emph{A Majority of Studies are Underpowered for Detecting Small
Effect Size Differences:}} Based on the previous observations we then
assessed whether the studies that we included are realistically powered
to identify small, medium, and large scale differences between case and
control. When assessing the power of each study at different effect
sizes the majority of studies for both adenoma and carcinoma have an
80\% power to detect a 30\% difference {[}Figure 7A \& B{]}. No single
study that was analyzed had the standard 80\% power to detect an effect
size difference that was equal to or below 10\% {[}Figure 7A \& B{]}. In
order to achieve adequate power for small effect sizes, studies would
need to recruit over 1000 individuals for each arm {[}Figure 7C{]}.

@@

\textbf{\emph{Assessing Important Random Forest Model Variables:}} The
genus based models collected the top 10 most important variables from
each training set and assessed how many times that genera showed up in
the top 10 across each study. The OTU based models recorded the medians
for each OTU across 100 different 80/20 splits of the data for each
study. The lowest classification for each OTU was obtained using the RDP
database and the number of times the specific classification occured in
the top 10 across studies was recorded. For the adenoma tissue genus and
OTU models there was only one matched and unmatched study and these
results were grouped together for the counting of the top 10.

@@

\textbf{Figure 6: Most Common Genera Across Full Community Stool Study
Models.} A) Common genera in the top 10 for adenoma Random Forest genus
models. B) Common genera in the top 10 for carcinoma Random Forest genus
models. C) Common genera in the top 10 for adenoma Random Forest OTU
models. D) Common genera in the top 10 for carcinoma Random Forest OTU
models.

\textbf{Figure 7: Power and Effect Size Analysis of Studies Included.}
A) Power based on effect size for studies with adenoma individuals. B)
Power based on effect size for studies with carcinoma individuals. C)
The estimated sample number needed for each arm of each study to detect
an effect size of 1-30\%. The dotted red lines in A) and B) represent a
power of 0.8.

@@

\textbf{Figure S7: Most Common Genera Across Full Community Tissue Study
Models.} A) Common genera in the top 10 for adenoma Random Forest genus
models. B) Common genera in the top 10 for unmatched carcinoma Random
Forest genus models. B) Common genera in the top 10 for matched
carcinoma Random Forest genus models. D) Common genera in the top 10 for
adenoma Random Forest OTU models. E) Common genera in the top 10 for
unmatched carcinoma Random Forest OTU models. F) Common genera in the
top 10 for matched carcinoma Random Forest OTU models.
